% FC 2027 paper — The Cost of Custody
% Target: 15pp Springer LNCS main body, unlimited references + appendix
% Double-blind submission

\documentclass[runningheads]{llncs}

% ===================== Packages =====================
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsmath,amssymb}
\usepackage{mathtools}
\usepackage{booktabs}
\usepackage{array}
\usepackage{tabularx}
\usepackage{multirow}
\usepackage[inline]{enumitem}
\usepackage{xcolor}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows.meta,positioning,calc,fit,backgrounds}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usepackage{siunitx}
\usepackage{url}
\usepackage{microtype}
\usepackage[bookmarksnumbered,hidelinks]{hyperref}

% ===================== Custom macros =====================
% Transaction size units
\newcommand{\vB}{\text{\,vB}}
\newcommand{\satvB}{\text{\,sat/vB}}

% Design / proposal names — roman, short OP_ prefix throughout. The
% full opcode names are glossed once on first use; thereafter we use
% these shorthands uniformly. For an actual opcode invocation in a
% code-like context (e.g.\ in an assertion or script listing), wrap
% locally in \texttt{...}.
\newcommand{\opctv}{OP\_CTV}
\newcommand{\opccv}{OP\_CCV}
\newcommand{\opvault}{OP\_VAULT}
\newcommand{\opvaultrecover}{OP\_VAULT\_RECOVER}
\newcommand{\opcat}{OP\_CAT}
\newcommand{\opcsfs}{OP\_CSFS}
\newcommand{\catcsfs}{\opcat{}+\opcsfs{}}

% Bitcoin concept names
\newcommand{\csv}{\texttt{OP\_CSV}}
\newcommand{\sighash}[1]{\texttt{SIGHASH\_#1}}

% Design-space axis labels --- canonical, see conference/AXES.md.
% Use these macros in every axis reference across FC paper and thesis;
% do not introduce synonyms (no "recovery-auth" where \axauth{} applies).
\newcommand{\axauth}{A1}          % Recovery authorisation  (formal: g)
\newcommand{\axrevault}{A2}       % Revault support         (formal: a)
\newcommand{\axfee}{A3}           % Fee-inclusion model     (formal: f)
\newcommand{\axwdbind}{A4}        % Withdrawal-destination binding (subcomp. of b)
\newcommand{\axrecbind}{A5}       % Recovery-destination binding   (subcomp. of b)
\newcommand{\axopsurf}{A6}        % Opcode surface          (new)
\newcommand{\axintro}{A7}         % Introspection primitive (new)

% Axis-with-formal-symbol aside (for proof-relevant passages).
\newcommand{\axauthf}{\axauth\,($g$)}
\newcommand{\axrevaultf}{\axrevault\,($a$)}
\newcommand{\axfeef}{\axfee\,($f$)}
\newcommand{\axbindf}{\axwdbind{}+\axrecbind\,($b$)}

% Attack-class labels --- canonical, see conference/AXES.md §4.
\newcommand{\classfee}{Z1}        % Fee-channel pinning
\newcommand{\classgrief}{Z2}      % Permissionless griefing
\newcommand{\classscale}{Z3}      % Per-UTXO recovery-cost scaling
\newcommand{\classhot}{Z4}        % Hot-key theft via destination substitution
\newcommand{\classcold}{Z5}       % Cold-key theft via unconstrained recovery
\newcommand{\classmode}{Z6}       % Mode/parameter bypass

% Draft-mode \todo — replaced or removed before submission
\newcommand{\todo}[2][]{\textbf{[TODO: #2]}}

% Measured-value placeholders. These render as [?] until the regtest
% harness produces the corresponding measurement (see
% vault-comparison/experiments/exp_watchtower_exhaustion.py Phase 6a).
% Replace each \measured{...} call with the numeric value before submission.
\newcommand{\measured}[1]{\textbf{[measured:#1]}}

% Batched-recovery constants — placeholders until Phase 6a runs on regtest
\newcommand{\overheadCCV}{54}
\newcommand{\perinputCCV}{68}
\newcommand{\overheadOPV}{154}
\newcommand{\perinputOPV}{92}

% ===================== Theorem environments =====================
% LNCS already defines theorem, definition, proposition, corollary, proof
% via \spnewtheorem in llncs.cls. No additional definitions needed.

% ===================== Paper metadata =====================
\title{The Cost of Custody}
\titlerunning{The Cost of Custody}

% Double-blind: authors anonymized until acceptance
\author{Anonymous Authors}
\authorrunning{Anonymous Authors}
\institute{Anonymous Institution\\
\email{anonymous@submission.invalid}}

\begin{document}

\maketitle

% ===================== Abstract =====================
\begin{abstract}
Bitcoin vaults encumber self-custodied funds with a time-delayed
spending path, allowing a compromised hot-key spend to be intercepted
and redirected to cold storage. Four script-extension proposals are
presently under review as candidates for enabling this pattern.
Practitioners must choose among them, yet no systematic cross-design
security analysis has been published.

We put forth a systematisation of the covenant-vault design space.
Our approach decomposes each design into four nearly-orthogonal
axes---a fee-payment model, an amount-flexibility policy, a
recovery-gating mechanism, and a recovery binding---each of which
deterministically opens a specific class of attack surface. We
formalise a unified defender-loss metric combining theft, griefing,
and dust-lockup outcomes, and identify a \emph{dominant corner} of
the 4D lattice: the feature combination at which all four surfaces
are closed simultaneously. No proposed Bitcoin Script extension
occupies this corner; each falls short in at least one axis, and
which axis it misses predicts the attack class that bites it.

We apply the framework to the four proposals, classify each in the
lattice, and validate the predicted vulnerability profiles with 15
regtest experiments and bounded model checking in Alloy (42
assertions across five formal models). We additionally measure
batched-recovery cost scaling and show that the safety ranking
between two proposals inverts depending on whether the watchtower
batches recovery---an operator-dependence that prior work has not
surfaced. A Simplicity vault on a Bitcoin-adjacent federated
sidechain serves as a cross-substrate reference point.

Our work shows that the cost of custody is not a single number but
a function of design choice, threat model, fee regime, and
watchtower strategy, and identifies a structurally-safer corner of
the design space that the current specification corpus has not yet
occupied.
\end{abstract}

% ===================== Sections =====================
\input{sections/01-introduction}
\input{sections/02-background}
\input{sections/03-impossibility}
\input{sections/04-empirical}
\input{sections/05-deployment}
\input{sections/06-related}
\input{sections/07-discussion}
\input{sections/08-conclusion}

% ===================== Bibliography =====================
\bibliographystyle{splncs04}
\bibliography{references}

% ===================== Appendix =====================
\appendix
\input{appendix-a-alloy}
\input{appendix-b-threat-matrix}
\input{appendix-c-measurements}
\input{appendix-d-simplicity}

\end{document}
