\chapter{Simplicity on Elements: The Unoccupied Optimal Corner}
\label{app:simplicity}

\section{Scope and Motivation}
\label{app:simplicity:scope}

Simplicity~\cite{OC17} is an \emph{alternative script primitive} --- a
typed combinator calculus with Coq-verified denotational semantics ---
deployed on Elements~\cite{DPNS16}, a Bitcoin-adjacent sidechain whose
blocks are produced by known functionaries rather than by proof-of-work
miners. Execution uses native-code \emph{jets} for known combinator
subtrees. Our Simplicity vault is excluded from the formal results of
Chapter~\ref{ch:security} and reported here as a cross-substrate
reference.

Chapter~\ref{ch:security}'s contribution is that the dominant corner
$D^\star$ is \emph{unoccupied on Bitcoin} among current proposals.
Simplicity provides a constructive existence proof that the corner is
occupiable on some substrate.

\section{Design-Space Position}
\label{app:simplicity:axes}

Our Simplicity vault on Elements supports per-output binding on both
trigger and recovery paths, atomic or bounded partial withdrawal,
cold-key recovery with an optional keyless post-timeout sweep, and
MAX\_FEE-capped in-value fee deduction. In the atomic configuration
with cold-key-only recovery, it occupies a strict-best position on
every axis where such a position exists in the taxonomy of
Section~\ref{se:design_space_axes}:

\begin{itemize}
  \item \axauth{} = \texttt{cold-key}: key-gated recovery via a
    Schnorr signature check. An optional construction exposes a
    keyless sweep to a pre-committed heir address once
    \texttt{jet::check\_lock\_height()} passes.
  \item \axrevault{} = \texttt{no} (atomic) or \texttt{yes-bounded}
    (bounded-partial with a compile-time \texttt{RESIDUAL\_FLOOR}
    covenant that structurally caps the resulting UTXO count).
  \item \axfee{} = \texttt{dynamic}: in-value fee deduction capped at a
    compile-time \texttt{MAX\_FEE} ceiling. The signer chooses the
    actual fee at broadcast. Elements carries an explicit fee output;
    there is no CPFP anchor surface.
  \item \axwdbind{} = \texttt{dedicated-jets-combinator}: the
    withdrawal destination is bound through
    \texttt{jet::output\_script\_hash(0)} and
    \texttt{jet::output\_amount(0)} at compile time, stronger than
    any of the Bitcoin proposals on the same axis.
  \item \axrecbind{} = \texttt{dedicated-jets-combinator}: the same
    jets bind the recovery output, eliminating class \classcold{}
    for the atomic cold-key-only configuration.
  \item \axopsurf{} = \texttt{single-mode}: a program's meaning is
    determined by its Merkle root; no \texttt{OP\_SUCCESSx}-style
    fallthrough surface.
  \item \axintro{} = \texttt{typed-combinator-introspection}: the
    fundamental operators are total, typed, and Coq-verified.
\end{itemize}

In the atomic configuration, our Simplicity vault is immune by
construction to classes \classfee{}, \classgrief{}, \classscale{},
\classhot{}, \classcold{}, and \classmode{} --- every attack class
identified in Chapter~\ref{ch:security}. It occupies the strict-best
position on \axauth{}, \axrevault{}, \axwdbind{}, \axrecbind{}, and
\axintro{} simultaneously; no Bitcoin proposal does. The
bounded-partial and keyless-post-timeout constructions admit
additional exposures documented in Section~\ref{app:simplicity:threats}
below.

\section{Lifecycle and Capabilities}
\label{app:simplicity:empirical}

The lifecycle measurement of $865\vB{}$ (deposit $269$, trigger $298$,
withdraw $298$, recover $286$), $49$--$135\%$ larger than the Bitcoin
designs of Table~\ref{tab:lifecycle}, is for the atomic configuration
with cold-key-only recovery. The overhead bundles Elements'
serialisation ($30$--$40$ bytes per output for confidential-transaction
fields even in non-confidential mode) and Simplicity's bit-oriented
DAG witness format. Elements functionaries do not operate a public
mempool, so class \classfee{} (fee-channel pinning) has no substrate
analogue under honest-federation assumptions.

The bounded-partial and keyless-post-timeout constructions share the
same covenant infrastructure and jets, differing only in the Simplicity
program selected at spend time. Integration tests cover lifecycle and
recovery for all three configurations; full lifecycle vsize
measurements for bounded-partial and keyless-post-timeout remain open.

The MAX\_FEE design deserves a note: the in-value deduction is capped
at a compile-time ceiling and the signer picks the actual fee at
broadcast. The cap keeps a compromised signer from burning the full
vault amount as fee.

\section{Threat Coverage}
\label{app:simplicity:threats}

The mapping below records, for a Simplicity vault on Elements, which
threat models are structurally addressed, which are not applicable
and for what reason, and which are addressed with a trade-off.

\begin{table}[h]
\centering
\footnotesize
\begin{tabular}{@{}lll@{}}
\toprule
Threat & Simplicity & Notes \\
\midrule
TM1 fee pinning            & N/A (substrate)     & Elements fee outputs; no CPFP anchor                                         \\
TM2 recovery griefing      & covered             & cold-key recovery; keyless post-timeout sweep, if configured, widens surface \\
TM3 timelock race          & covered             & CSV via \texttt{jet::check\_lock\_distance()}                                \\
TM4 / \classscale{}        & atomic: covered;     & residual-floor covenant structurally caps UTXO count                        \\
                           & bounded-partial: bounded &                                                                         \\
TM5 trigger-xpub theft     & covered             & two-key model                                                                \\
TM6 dual-key compromise    & N/A (scope)         & Kerckhoffs out of scope                                                      \\
TM7 address reuse          & covered             & fresh plan per vault                                                         \\
TM8 CCV mode bypass        & N/A (design)        & single-mode on A6                                                            \\
TM9 hot-key redirection    & covered             & per-output binding prevents redirection                                      \\
TM10 witness tampering     & covered             & typed-combinator witness check                                               \\
TM11 cold-key compromise   & covered (forced recovery) & keyless post-timeout sweep, if configured, introduces a heir-timing trade-off \\
\bottomrule
\end{tabular}
\caption{Threat coverage for our Simplicity vault on Elements. Cells
labelled \emph{bounded} (TM4 under bounded-partial) and
\emph{trade-off} (TM2 and TM11 under the keyless post-timeout
construction) mark design-space points that are not captured by the
binary covered/uncovered split used in Chapter~\ref{ch:security}.}
\label{tab:simplicity-threats}
\end{table}

\section{Foundations and Substrate}
\label{app:simplicity:foundations}

Simplicity's Coq-verified denotational semantics give stronger
guarantees than Alloy bounded model checking of the Bitcoin Script
extensions can produce --- a Simplicity program's meaning is
determined by its Merkle root, under a formal semantics that has
been mechanically verified. Simplicity on Liquid is live today under
\emph{federation trust}; the four Bitcoin Script extensions remain
under \emph{Bitcoin consensus review}.

The substrate choice is a security-relevant decision alongside the
per-covenant analysis of Chapter~\ref{ch:security}: a vault on
Simplicity inherits the federation's liveness and censorship
properties; a vault on Bitcoin inherits proof-of-work's. The
dominant corner being occupiable on Simplicity is therefore not a
turn-key prescription to migrate vaults to Liquid; it is evidence
that the unoccupied status among current Bitcoin proposals reflects a
consensus-substrate constraint rather than a design impossibility.

\section{Future Work: Porting the Optimal Corner to Bitcoin}
\label{app:simplicity:future}

A Simplicity-style vault on Bitcoin would require Simplicity's
taproot leaf type, which is specified in a draft BIP but not under
active Bitcoin consensus review. Closing the Simplicity-to-Bitcoin
gap depends on that soft-fork path. An alternative that preserves
Bitcoin's consensus substrate is a CAT+CSFS-based bound-recovery
construction (Poelstra's sketch~\cite{Poe21}) that moves CAT+CSFS
to \axrecbind{} = \texttt{signature-dual-bind}; this would eliminate
class \classcold{} while retaining the hot-key-theft immunity of
class \classhot{}, placing CAT+CSFS at the dominant corner of the
Bitcoin sublattice. Neither path is available as of thesis
submission; we leave both as open problems. The bounded-partial
construction's residual-floor covenant is a further example of a
structural bound; the question of how close such a bound can come to
full class-\classscale{} immunity is left open.
