% ══════════════════════════════════════════════════════════════
%  The Cost of Custody: Security Tradeoffs and Fee Dynamics in Bitcoin Covenant Vaults
%  Master's Thesis — Virginia Tech ETD Format
%  Compile: xelatex thesis && bibtex thesis && xelatex thesis && xelatex thesis
% ══════════════════════════════════════════════════════════════

\documentclass[doublespace,nopageskip]{VTthesis}

% ── Additional packages ───────────────────────────────────────
\usepackage{booktabs}       % Professional tables
\usepackage{multirow}       % Multi-row table cells
\usepackage{siunitx}        % Number formatting (vB, sats)
\usepackage{pgfplots}       % Data plots
\pgfplotsset{compat=1.18}
\usepackage{listings}       % Code listings (Alloy, Simplicity)
\usepackage{xcolor}         % Colors for listings and TikZ
\usepackage{subcaption}     % Subfigures
\usepackage{rotating}       % Landscape tables
\usepackage{pdflscape}      % Landscape pages
\usepackage{longtable}      % Multi-page tables
\usepackage{url}            % URL formatting
\usepackage{amsmath,amssymb,amsthm}

% ── Fix vertical spacing ─────────────────────────────────────
\raggedbottom                % Prevent LaTeX from stretching vertical glue to fill pages

% ── Fix overfull hbox from long texttt ────────────────────────
\tolerance=1000              % Allow slightly looser line breaking
\emergencystretch=1em        % Extra stretch to avoid overfull boxes
\setlength{\emergencystretch}{2em}

% ── Listings style ────────────────────────────────────────────
\lstdefinestyle{alloy}{
  basicstyle=\ttfamily\small,
  keywordstyle=\bfseries,
  commentstyle=\itshape\color{gray},
  breaklines=true,
  frame=single,
  numbers=left,
  numberstyle=\tiny\color{gray},
}

\lstdefinestyle{simf}{
  basicstyle=\ttfamily\small,
  keywordstyle=\bfseries\color{blue!70!black},
  commentstyle=\itshape\color{gray},
  stringstyle=\color{red!70!black},
  breaklines=true,
  frame=single,
  numbers=left,
  numberstyle=\tiny\color{gray},
  morekeywords={fn,let,match,Left,Right,assert,param,witness,jet},
}

% ── TikZ libraries ────────────────────────────────────────────
\usetikzlibrary{shapes,arrows.meta,positioning,calc,fit,backgrounds}

% ── Custom commands ───────────────────────────────────────────
\newcommand{\vB}{\,\text{vB}}               % virtual bytes
\newcommand{\satvB}{\,\text{sat/vB}}        % fee rate units
\newcommand{\opctv}{\texttt{OP\_CTV}}
\newcommand{\opccv}{\texttt{OP\_CCV}}
\newcommand{\opvault}{\texttt{OP\_VAULT}}
\newcommand{\catcsfs}{\texttt{CAT+CSFS}}
\newcommand{\opcat}{\texttt{OP\_CAT}}
\newcommand{\opcsfs}{\texttt{OP\_CHECKSIGFROMSTACK}}
\newcommand{\sighash}[1]{\texttt{SIGHASH\_#1}}

% Theorem environments provided by VTthesis.cls (theorem, corollary, definition)

% ── Metadata ──────────────────────────────────────────────────
\title{The Cost of Custody}

\keywords{Bitcoin, covenants, vaults, formal verification, transaction analysis}

\author{Praneeth Gunasekaran}

\program{Computer Science}

\degree{Master of Science}

\submitdate{April 29, 2026}

\principaladvisor{Shinichiro Matsuo}
\coadvisor{Wenjing Lou}
\firstreader{Na Meng}

\dedication{To my parents and family, for their unconditional love and support,
and to the Bitcoin open-source community, whose relentless pursuit of
self-sovereign custody inspired this work.}

\acknowledge{I would like to express my sincere gratitude to my advisor,
Dr.\ Shinichiro Matsuo, for his guidance, encouragement, and support throughout
this research. I am also grateful to my co-advisor, Dr.\ Wenjing Lou, and
committee member, Dr.\ Na Meng, for their valuable feedback and time reviewing
this thesis.

I owe a debt to the Bitcoin covenant developers whose work made this research
possible. Jeremy Rubin provided feedback on the overall research direction and
helped shape the comparative methodology. James O'Beirne gave detailed feedback
on the OP\_VAULT experiments, including insights on reducing fee overhead in the
vault lifecycle. Salvatore Ingala's OP\_CHECKCONTRACTVERIFY design and pymatt
framework were essential to the CCV experiments. Andrew Poelstra's CAT and
Schnorr Tricks work laid the foundation for the CAT+CSFS vault. The Blockstream
Research team---particularly Russell O'Connor (Simplicity) and the Simplex SDK
developers---helped narrow down the research ideas and provided the language and
tooling for the Simplicity vault implementation.

I would also like to thank the Commonwealth Cyber Initiative for supporting
this work.

Finally, I thank my friends and family for their patience, support, and
encouragement throughout my graduate studies.}

\abstract{%
Bitcoin covenant proposals enable self-custody vaults that enforce spending
conditions beyond simple signature checks. Four proposals---\opctv{} (BIP-119),
\opccv{} (BIP-443), \opvault{} (BIP-345), and \catcsfs{} (BIP-347 + BIP-348)---%
represent fundamentally different design philosophies, yet no prior work compares
them empirically. We build all four vault implementations, run 15~experiments on
Bitcoin regtest, and extend the comparison to a fifth vault built with the
Simplicity language on Elements regtest. We establish two results that constrain the
design space: (1)~no covenant vault achieves both permissionless recovery and
griefing resistance (Proposition~1), and (2)~the relative security ordering of vault
designs inverts depending on the fee environment (Theorem~1), with a crossover
near 50~sat/vB for a 0.5~BTC vault. We correct prior hand-estimates of OP\_VAULT
transaction sizes by 46\%, measure CCV's watchtower splitting capacity at
${\sim}6{,}172$~splits per block (80\% more than previously estimated), and
provide the first parameterized attack cost functions across fee environments.
Alloy bounded model checking verifies fund conservation, recovery liveness, and
no-extraction properties across all five designs. The framework, measurements,
and formal models are open-source.%
}

\abstractgenaud{%
Bitcoin users who store their own cryptocurrency face a fundamental problem:
if someone steals their private key, the funds are gone. Covenant vaults are
a proposed solution---special Bitcoin transactions that enforce a time delay
before funds can be moved, giving the owner a window to recover stolen coins.
Four different vault designs have been proposed, each with different tradeoffs.
We built all four, tested them side-by-side, and discovered that no single
design is safest in all situations. The best choice depends on transaction
fees: when fees are low, designs that allow partial withdrawals are safer;
when fees are high, simpler designs that resist fee-based attacks become
preferable. We also found that existing estimates of transaction costs were
significantly wrong---one design is 46\% more expensive than previously thought.
Our measurements and testing framework are publicly available.%
}

% ══════════════════════════════════════════════════════════════
\begin{document}

% iThenticate version: title page excluded per VT Graduate School instructions
\frontmatter
%\maketitle  % EXCLUDED for iThenticate
\tableofcontents
\listoffigures
\listoftables

% ── List of Abbreviations ─────────────────────────────────────
\chapter*{List of Abbreviations}
\addcontentsline{toc}{chapter}{List of Abbreviations}
\begin{tabbing}
\hspace{3cm} \= \kill
BIP     \> Bitcoin Improvement Proposal \\
CAT     \> OP\_CAT (BIP-347), stack concatenation opcode \\
CCV     \> OP\_CHECKCONTRACTVERIFY (BIP-443) \\
CLTV    \> OP\_CHECKLOCKTIMEVERIFY, absolute timelock \\
CPFP    \> Child Pays For Parent, fee bumping mechanism \\
CSFS    \> OP\_CHECKSIGFROMSTACK (BIP-348) \\
CSV     \> OP\_CHECKSEQUENCEVERIFY, relative timelock \\
CTV     \> OP\_CHECKTEMPLATEVERIFY (BIP-119) \\
DAG     \> Directed Acyclic Graph \\
HTLC    \> Hash Time-Locked Contract \\
P2TR    \> Pay-to-Taproot (Taproot output type) \\
P2WPKH  \> Pay-to-Witness-Public-Key-Hash \\
P2WSH   \> Pay-to-Witness-Script-Hash \\
RBF     \> Replace-By-Fee \\
TRUC    \> Topologically Restricted Until Confirmation (v3 transactions) \\
UTXO    \> Unspent Transaction Output \\
vB      \> Virtual bytes (segwit-discounted transaction size) \\
WU      \> Weight units (1 vB = 4 WU)
\end{tabbing}

% ══════════════════════════════════════════════════════════════
\mainmatter

\include{chapters/ch1_introduction}
\include{chapters/ch2_background}
\include{chapters/ch3_methodology}
\include{chapters/ch4_results}
\include{chapters/ch5_security}
\include{chapters/ch6_verification}
\include{chapters/ch7_discussion}
\include{chapters/ch8_related_work}
\include{chapters/ch9_conclusion}

% ── Bibliography ──────────────────────────────────────────────
% EXCLUDED for iThenticate per VT Graduate School instructions
%\bibliography{references}
%\bibliographystyle{plainnat}

% ── Appendix ──────────────────────────────────────────────────
\appendix
\begin{appendices}
  \include{chapters/app_a_alloy}
\end{appendices}

\end{document}
