% ══════════════════════════════════════════════════════════════
%  The Cost of Custody: Security Tradeoffs and Fee Dynamics in Bitcoin Covenant Vaults
%  Master's Thesis — Virginia Tech ETD Format
%  Compile: xelatex thesis && bibtex thesis && xelatex thesis && xelatex thesis
% ══════════════════════════════════════════════════════════════

\documentclass[doublespace,nopageskip]{VTthesis}

% ── Additional packages ───────────────────────────────────────
\usepackage{booktabs}       % Professional tables
\usepackage{multirow}       % Multi-row table cells
\usepackage{siunitx}        % Number formatting (vB, sats)
\usepackage{pgfplots}       % Data plots
\pgfplotsset{compat=1.18}
\usepackage{listings}       % Code listings (Alloy, Simplicity)
\usepackage{xcolor}         % Colors for listings and TikZ
\usepackage{subcaption}     % Subfigures
\usepackage{rotating}       % Landscape tables
\usepackage{pdflscape}      % Landscape pages
\usepackage{longtable}      % Multi-page tables
\usepackage{url}            % URL formatting
\usepackage{amsmath,amssymb,amsthm}
\usepackage{enumitem}       % Customisable list environments (description[leftmargin=...])

% ── Fix vertical spacing ─────────────────────────────────────
\raggedbottom                % Prevent LaTeX from stretching vertical glue to fill pages

% ── Fix overfull hbox from long texttt ────────────────────────
\tolerance=1000              % Allow slightly looser line breaking
\emergencystretch=1em        % Extra stretch to avoid overfull boxes
\setlength{\emergencystretch}{2em}

% ── Listings style ────────────────────────────────────────────
\lstdefinestyle{alloy}{
  basicstyle=\ttfamily\small,
  keywordstyle=\bfseries,
  commentstyle=\itshape\color{gray},
  breaklines=true,
  frame=single,
  numbers=left,
  numberstyle=\tiny\color{gray},
}

\lstdefinestyle{simf}{
  basicstyle=\ttfamily\small,
  keywordstyle=\bfseries\color{blue!70!black},
  commentstyle=\itshape\color{gray},
  stringstyle=\color{red!70!black},
  breaklines=true,
  frame=single,
  numbers=left,
  numberstyle=\tiny\color{gray},
  morekeywords={fn,let,match,Left,Right,assert,param,witness,jet},
}

% ── TikZ libraries ────────────────────────────────────────────
\usetikzlibrary{shapes,arrows.meta,positioning,calc,fit,backgrounds}

% ── Custom commands ───────────────────────────────────────────
\newcommand{\vB}{\,\text{vB}}               % virtual bytes
\newcommand{\satvB}{\,\text{sat/vB}}        % fee rate units
\newcommand{\opctv}{\texttt{OP\_CTV}}
\newcommand{\opccv}{\texttt{OP\_CCV}}
\newcommand{\opvault}{\texttt{OP\_VAULT}}
\newcommand{\catcsfs}{\texttt{CAT+CSFS}}
\newcommand{\opcat}{\texttt{OP\_CAT}}
\newcommand{\opcsfs}{\texttt{OP\_CHECKSIGFROMSTACK}}
\newcommand{\sighash}[1]{\texttt{SIGHASH\_#1}}

% Design-space axis labels --- canonical, shared with FC 2027 paper.
% See conference/AXES.md for the canonical definitions.
\newcommand{\axauth}{A1}          % Recovery authorisation  (formal: g)
\newcommand{\axrevault}{A2}       % Revault support         (formal: a)
\newcommand{\axfee}{A3}           % Fee-inclusion model     (formal: f)
\newcommand{\axwdbind}{A4}        % Withdrawal-destination binding (subcomp. of b)
\newcommand{\axrecbind}{A5}       % Recovery-destination binding   (subcomp. of b)
\newcommand{\axopsurf}{A6}        % Opcode surface          (new)
\newcommand{\axintro}{A7}         % Introspection primitive (new)

% Axis-with-formal-symbol aside (for proof-relevant passages).
\newcommand{\axauthf}{\axauth\,($g$)}
\newcommand{\axrevaultf}{\axrevault\,($a$)}
\newcommand{\axfeef}{\axfee\,($f$)}
\newcommand{\axbindf}{\axwdbind{}+\axrecbind\,($b$)}

% Attack-class labels --- canonical.
\newcommand{\classfee}{Z1}        % Fee-channel pinning
\newcommand{\classgrief}{Z2}      % Permissionless griefing
\newcommand{\classscale}{Z3}      % Per-UTXO recovery-cost scaling
\newcommand{\classhot}{Z4}        % Hot-key theft via destination substitution
\newcommand{\classcold}{Z5}       % Cold-key theft via unconstrained recovery
\newcommand{\classmode}{Z6}       % Mode/parameter bypass

% Theorem environments provided by VTthesis.cls (theorem, corollary, definition)

% ── Metadata ──────────────────────────────────────────────────
\title{The Cost of Custody}

\keywords{Bitcoin, Script extensions, vault custody, covenants, formal verification, fee dynamics}

\author{Praneeth Gunasekaran}

\program{Computer Science}

\degree{Master of Science}

\submitdate{April 29, 2026}

\principaladvisor{Shinichiro Matsuo}
\coadvisor{Wenjing Lou}
\firstreader{Na Meng}

\dedication{To my parents and family, for their unconditional love and support,
and to the Bitcoin open-source community, whose relentless pursuit of
self-sovereign custody inspired this work.}

\acknowledge{I would like to express my sincere gratitude to my advisor,
Dr.\ Shinichiro Matsuo, for his guidance, encouragement, and support throughout
this research. I am also grateful to my co-advisor, Dr.\ Wenjing Lou, and
committee member, Dr.\ Na Meng, for their valuable feedback and time reviewing
this thesis.

I owe a debt to the Bitcoin covenant developers whose work made this research
possible. Jeremy Rubin provided feedback on the overall research direction and
helped shape the comparative methodology. James O'Beirne gave detailed feedback
on the OP\_VAULT experiments, including insights on reducing fee overhead in the
vault lifecycle. Salvatore Ingala's OP\_CHECKCONTRACTVERIFY design and pymatt
framework were essential to the CCV experiments. Andrew Poelstra's CAT and
Schnorr Tricks work laid the foundation for the CAT+CSFS vault. The Blockstream
Research team---particularly Russell O'Connor (Simplicity) and the Simplex SDK
developers---helped narrow down the research ideas and provided the language and
tooling for the Simplicity vault implementation.

I would also like to thank the Commonwealth Cyber Initiative for supporting
this work.

Finally, I thank my friends and family for their patience, support, and
encouragement throughout my graduate studies.}

\abstract{%
Four Bitcoin Script extension proposals---\opctv{} (BIP-119),
\opvault{} (BIP-345), \opccv{} (BIP-443), and the \opcat{} +
\opcsfs{} composition (BIP-347 + BIP-348)---enable self-custody vaults
that enforce spending conditions beyond simple signature checks, but
no prior work compares them empirically. We build all four vault
implementations, run 15~experiments on Bitcoin regtest, and additionally
implement Simplicity on Elements regtest as a cross-substrate reference
point. We formalise a defender-loss metric~$\mathcal{L}$ and, threat by
threat, characterise how~$\mathcal{L}$ varies with the fee rate across
designs. Two results constrain the design space: (1)~a
Griefing--Safety Incompatibility (no vault simultaneously achieves
permissionless recovery and griefing resistance), and (2)~a No
Pareto-Optimal Design theorem (for every design, there exists a threat
model and fee rate at which another design has strictly lower defender
loss). We measure OP\_VAULT and CCV watchtower splitting capacity
(3{,}427 and 6{,}172 splits/block respectively), show that
defender-side batched recovery---spec-level in BIP-345 §``Batching''
and BIP-443 aggregation---approximately doubles the dust-threshold
fee rate and flips the CCV/OP\_VAULT ordering under watchtower
exhaustion, and provide the first parameterised attack cost functions
across fee environments. Alloy bounded model checking verifies fund
conservation, recovery liveness, and no-extraction properties.
Simplicity's federated-sidechain threat model is treated separately.
The framework, measurements, and formal models are open-source.%
}

\abstractgenaud{%
Bitcoin users who store their own cryptocurrency face a fundamental problem:
if someone steals their private key, the funds are gone. \emph{Vault}
protocols are a proposed solution---specially-scripted Bitcoin
transactions that enforce a time delay before funds can be moved,
giving the owner a window to recover stolen coins. Four different
vault designs have been proposed for Bitcoin, each with different
tradeoffs. We built all four, tested them side-by-side, and discovered
that no single design is safest against every threat: the right choice
depends on which attack you most fear and on the transaction fee
environment. We also found that a batching optimisation built into
two of the designs roughly doubles the fee rate at which a common
denial-of-service attack becomes practical, and unexpectedly reverses
the relative safety of those two designs. We additionally studied
Simplicity, a fundamentally different scripting language deployed on
a Bitcoin-adjacent sidechain, as a point of comparison. Our
measurements and testing framework are publicly available.%
}

% ══════════════════════════════════════════════════════════════
\begin{document}

\frontmatter
\maketitle
\tableofcontents
\listoffigures
\listoftables

% ── List of Abbreviations ─────────────────────────────────────
\chapter*{List of Abbreviations}
\addcontentsline{toc}{chapter}{List of Abbreviations}
\begin{tabbing}
\hspace{3cm} \= \kill
BIP     \> Bitcoin Improvement Proposal \\
CAT     \> OP\_CAT (BIP-347), stack concatenation opcode \\
CCV     \> OP\_CHECKCONTRACTVERIFY (BIP-443) \\
CLTV    \> OP\_CHECKLOCKTIMEVERIFY, absolute timelock \\
CPFP    \> Child Pays For Parent, fee bumping mechanism \\
CSFS    \> OP\_CHECKSIGFROMSTACK (BIP-348) \\
CSV     \> OP\_CHECKSEQUENCEVERIFY, relative timelock \\
CTV     \> OP\_CHECKTEMPLATEVERIFY (BIP-119) \\
DAG     \> Directed Acyclic Graph \\
HTLC    \> Hash Time-Locked Contract \\
P2TR    \> Pay-to-Taproot (Taproot output type) \\
P2WPKH  \> Pay-to-Witness-Public-Key-Hash \\
P2WSH   \> Pay-to-Witness-Script-Hash \\
RBF     \> Replace-By-Fee \\
TRUC    \> Topologically Restricted Until Confirmation (v3 transactions) \\
UTXO    \> Unspent Transaction Output \\
vB      \> Virtual bytes (segwit-discounted transaction size) \\
WU      \> Weight units (1 vB = 4 WU)
\end{tabbing}

% ══════════════════════════════════════════════════════════════
\mainmatter

\include{chapters/ch1_introduction}
\include{chapters/ch2_background}
\include{chapters/ch3_methodology}
\include{chapters/ch4_results}
\include{chapters/ch5_security}
\include{chapters/ch6_verification}
\include{chapters/ch7_discussion}
\include{chapters/ch8_related_work}
\include{chapters/ch9_conclusion}

% ── Bibliography ──────────────────────────────────────────────
\bibliography{references}
\bibliographystyle{plainnat}

% ── Appendix ──────────────────────────────────────────────────
\appendix
\begin{appendices}
  \include{chapters/app_a_alloy}
  \include{chapters/app_c_simplicity}
  \include{chapters/app_d_rationality}
\end{appendices}

\end{document}
